---
title: "Untitled"
output: 
  beamer_presentation:
classoption: "aspectratio=169"    
theme: "Singapore"
#colortheme: "dolphin"
fonttheme: "structurebold"
slide_level: 3
keep_tex: false
header-includes:
 - \usepackage{tcolorbox}
 - \tcbuselibrary{fitting}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(wooldridge)
library(stargazer)
```

## outline
\tableofcontents[hideallsubsections]

# Section 1
## Looking Straight at a Distribution

- The distribution of a variable contains *everything we know* about that variable from empirical observation
\pause
- Any description we make will be a *summary* of that distribution
- So we may as well look at it directly!

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.


## Distributions of Kinds of Variables

- There are two main kinds of variables for which the distributions look different: discrete and continuous
- Discrete variables take a finite set of values: left-handed, right-handed, ambidextrous. Or "lives in Seattle" vs. "Doesn't" or "Number of kids"
- Continuous variables take any value: income, height, KwH of electricity used each day
- (Sometimes, "ordinal" discrete variables with many values are treated as continuous for simplicity)


# Section 4
## Slide with Plot




```{r, echo = TRUE, eval=FALSE}
library(wooldridge)
data("wage1")
wageModel <- lm(lwage ~ educ + exper + tenure, data = wage1)
summary(wageModel)

```

---

```{r, echo = FALSE, eval=TRUE}


data("wage1")
wageModel <- lm(lwage ~ educ + exper + tenure, data = wage1)
stargazer(wageModel,type = "latex", out = "models.tex", title="first table")
```



## Results

\begin{tcolorbox}[fit,width=\textwidth,height=.9\textheight,size=minimal,colback=white,fit algorithm=fontsize,colframe=white]
  \input{models.tex}
\end{tcolorbox}


